---
title: "pL2003_OriginalHomeworkCode_04"
output: html_document
date: "2025-03-15"
---


```{r}
# Load required libraries
library(ggplot2)
library(dplyr)
```


```{r}
#p1, n1, p0 no default
#p2, n2 NULL default 
#alternative- "two sided" 
#conf lvl= 0.95

Z.prop.test <- function(p1, n1, p2 = NULL, n2 = NULL, p0, alternative = "two.sided", conf.level = 0.95) {
  
# Step 1: Rules of thumb 
if (n1 * p1 <= 5 || n1 * (1 - p1) <= 5) {warning("Normal approximation may not be valid: n1 * p1 or n1 * (1 - p1) <= 5.")} #warning message if it is not greater than 5 
  
if (!is.null(p2) & !is.null(n2)) {if (n2 * p2 <= 5 || n2 * (1 - p2) <= 5) {warning("Normal approximation may not be valid: n2 * p2 or n2 * (1 - p2) <= 5.")}}
  
# Step 2: One-sample test
if (is.null(p2) || is.null(n2)) {se <- sqrt(p0 * (1 - p0) / n1)  # Standard error
Z <- (p1 - p0) / se  # Z-score
estimate <- p1  # Estimate is just p1
} 

# Step 3: Two-sample test
else {pooled_p <- (p1 * n1 + p2 * n2) / (n1 + n2)  # Pooled proportion
se <- sqrt(pooled_p * (1 - pooled_p) * (1/n1 + 1/n2))  # Standard error
Z <- (p1 - p2) / se  # Z-score
estimate <- p1 - p2  # Estimate is the difference in proportions
}
  
# Step 4: P-value calculation
P <- if (alternative == "two.sided") {2 * (1 - pnorm(abs(Z)))} 
else if (alternative == "less") {pnorm(Z)} else if (alternative == "greater") {1 - pnorm(Z)} 
else {stop("Invalid alternative hypothesis. Use 'two.sided', 'less', or 'greater'.")}
  
# Step 5: Confidence interval
Z_critical <- qnorm(1 - (1 - conf.level) / 2)
margin <- Z_critical * se
CI <- c(estimate - margin, estimate + margin)
  
# Step 6: Return results
return(list(Z = Z, P = P, CI = CI))}
  
```
## Jon's Comment: "Your set up looks good. Very neat and readable. Nothing else to add."
## yz: I believe that using indents in the code would make it easier to track where a specific 
## if statement ends and improve the overall readability of the code flow. That said, the function 
## itself appears to be accurate and works well! I also really appreciated the clear and helpful comments
## throughout the code.
```{r, error=TRUE}

# Load the dataset 
data <- read.csv ("C:/Users/yunhu/Documents/BI588/AN588_Malfunction_yunhuaz/KamilarAndCooperData.csv") 

# View the structure of the dataset
str(data)

# Fit the first regression model (without log transformation)
model1 <- lm(MaxLongevity_m ~ Brain_Size_Species_Mean, data = data)

# Fit the second regression model (with log transformation)
model2 <- lm(log(MaxLongevity_m) ~ log(Brain_Size_Species_Mean), data = data)

# View the regression summaries
summary(model1)
summary(model2)

### YZ: From lines 79 - 91 I believe your missing "+" signs
# Scatterplot with regression line
plot1 <- ggplot(data, aes(x = Brain_Size_Species_Mean, y = MaxLongevity_m)) 

#dot colors 
plot1 <- plot1 + geom_point(color = "blue")  

#linear regression line, red, no shaded region around regression line to show se
plot1 <- plot1 + geom_smooth(method = "lm", color = "red", se = FALSE)  

 #labels for plot
plot1 <- plot1+ labs(title = "Longevity vs Brain Size", x = "Brain Size (grams)", y = "Longevity (months)") + theme_minimal()   #labels for plot

# Display the plot
plot1

## YZ: Missing "+" from lines 96-105 for the ggplots

plot2 <- ggplot(data, aes(x = log(Brain_Size_Species_Mean), y = log(MaxLongevity_m))) 

#dot colors 
plot2 <- plot2+ geom_point(color = "blue") 

#linear regression line, red, no shaded region around regression line to show se
plot2 <- plot2+ geom_smooth(method = "lm", color = "red", se = FALSE) 

#labels for plot
plot2 <- plot2+ labs(title = "Log-Log Relationship: Longevity vs Brain Size", x = "Log(Brain Size)", y = "Log(Longevity)") + theme_minimal()

# Display the plot
plot2

#calculate the confidence intervals for the parameters (coefficients) of lm
confint(model1, level = 0.90)  # 90% CI for Model 1
confint(model2, level = 0.90)  # 90% CI for Model 2

#Check for missing values and infinite values before analysis or modeling to prevent erros 
sum(is.na(data$Brain_Size_Species_Mean))  # Count NAs
sum(is.infinite(data$Brain_Size_Species_Mean))  # Count Inf values

#remove na and inf values 
data <- data %>% filter(!is.na(Brain_Size_Species_Mean) & is.finite(Brain_Size_Species_Mean))

### YZ: I didn't even realize there were NA values! I think it was very smart to remove them!

# Create prediction data
new_data <- data.frame(Brain_Size_Species_Mean = seq(min(data$Brain_Size_Species_Mean), max(data$Brain_Size_Species_Mean), length.out = 100))

#generate sequence with values between min and max to plot  

predictions <- predict(model1, newdata = new_data, interval = "prediction", level = 0.90) #generate generate prediction intervals for values of y at each x

confidence <- predict(model1, newdata = new_data, interval = "confidence", level = 0.90) #generate confidence intervals around our predicted mean value for y values

# Combine the predictions and confidence intervals into a new data frame
df <- cbind(new_data, predictions, confidence[,2:3])
# Rename the columns
names(df) <- c("x", "y", "yhat", "CIfit", "CIlwr", "CIupr")
# View the resulting data frame
head(df)

### YZ: Really liked that you created a new dataframe to incorporate all the values for easier reading!

# Plot with CI and PI bands
ggplot() +  
# Scatterplot of the data
geom_point(data = data, aes(x = Brain_Size_Species_Mean, y = MaxLongevity_m), color = "blue") +  
#Regression line 
geom_line(data = plot_data, aes(x = Brain_Size_Species_Mean, y = fit), color = "red") +  
#Confidence Interval Bounds (blue)
geom_line(data = plot_data, aes(x= Brain_Size_Species_Mean, y= lwr_conf), color= "blue")+
geom_line(data = plot_data, aes(x = Brain_Size_Species_Mean, y = upr_conf), color = "blue")+
# Prediction Interval (Green)
geom_line(data = plot_data, aes(x = Brain_Size_Species_Mean, y = lwr_pred), color = "green") +
geom_line(data = plot_data, aes(x = Brain_Size_Species_Mean, y = upr_pred), color = "green")

## Jon's Comment: "what's plot_data? Is that meant for new_data or your df?"
## YZ: I recieved the same error! I believe it should be df since its a dataframe with
## all your CI and PI intervals.


# Labels and theme
labs(title = "Regression with Confidence & Prediction Intervals", x = "Brain Size (grams)", y = "Longevity (months)") +  
theme_minimal()

### YZ: Lines 142-15, ggplot syntax!

# Create new data point for prediction
new_species <- data.frame(Brain_Size_Species_Mean = 800)

# Get prediction interval
predict(model1, new_species, interval = "prediction", level = 0.90)


## YZ: Noticed that you worked on model 1 to generate the predictions
## I think making one for model 2 can provide a good comparison between the models!
```
Producing the point estimate- The point estimate longevity for a species with 800g brain size is 1223.345 with a 90% prediction interval of the true longevity lying between 1021.805 and 1424.884


Comparing the models 
Using summary() for the Linear and Log-Log model, the log-log model shows a better fit for the data 
The residuals for log-log model is smaller and less spread out indicating that the transformation helped improve the fit of the model. 

The Log-Log model also has a higher R^2 value compared to the linear model (49.28%). For model 1, 49.28% of the variation in longevity is explained by brain size, but the remaining is not. For model 2, 57.8% is explained. 

The residual standard error for log-log is also much smaller than the linear model, indicating that the predictions are more precise. 

I would trust the model to predict the observations accurately, but since 800 is much larger than the other values from the data, I would be more careful when predicting.  

## Jon's Comment: "For the most part, everything looks good, except for the Scatterplot Data. It might have been an error there, but I wasn't able to run it."

## YZ: Great job on the code! I think breaking the code into small chunks and putting indents
## would improve readability and makes it easier to inspect outputs! I noticed a few syntax issues 
## in ggplot code that can be fixed. Don't forget to interpret the slop of the regression lines!
## Your explanatinos are clear, and with these tweeks, I believe you anaylsis will be even stronger!